### 2017-10-18 Draft2/Appending.Dynamic.tex
Add constructor code
Algorithm for Append
Initial Grow impl
Rename some things and remove some brackets

\begin{description}
	\item[initial capacity] The capacity of a dynamic a array when one item is appended to it. I denote this $\initcapacity$.\\
	(The capacity of any dynamic array with size zero should be zero.)
	\item[growth factor] The factor by which the current capacity is multiplied to get the new capacity when $\listname$ is resized. I denote this with $\growthfactor$.
\end{description}

% impl.

@@

Before I analyze the time complexity of [append], I consider a different method for measuring its cost. Suppose I start with an empty list and $\varnitems$ elements are appended. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $\varnitems$ appends, and denote it $\nwritesfn(n)$.

In the above code, one write is performed after the conditional, so it is apparent that $\nwritesfn(\varnitems) \geq \varnitems$ after $n$ appends. However, [resize] also does some writing, so in order to find a precise formula for $\nwritesfn(\varnitems)$ I need to analyze when [resize] is called. Consider the following lemma:

@@

[proof]
The capacity starts out at zero and becomes $\initcapacity$ once one element is appended. From then on, it can only grow by a factor of $\growthfactor$ via [resize], so if $\growthfactor^i\initcapacity$ is the current capacity then $\growthfactor^{i + 1}\initcapacity$ must be the next capacity. By induction, the rest of the sequence is $$\left\{ \growthfactor^i\initcapacity \right\}_{i = 1}^k$$ for some $k$. Since the capacity is only as big as is needs to be, namely greater than or equal to $\varnitems$, $k$ is the smallest integer for which $\growthfactor^k\initcapacity \geq \varnitems$. Using this property, I now derive $k$.

@@

[make lemma]
Now, consider that to grow by $\growthfactor$ [resize] must be called. So $\growthfactor^k\initcapacity$ being present in $S_C$ means [resize] was called at size $\growthfactor^{k - 1}\initcapacity$ for $k \geq 1$. Then [resize] was called at the sizes $$S_R = 0,\ \initcapacity,\ \growthfactor\initcapacity,\ \growthfactor^2\initcapacity,\ \ldots\ \growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil - 1}\initcapacity$$

Each time [resize] is called, $i$ items are copied where $i$ is the current size of the dynamic array. Then the total number of items copied for $\varnitems$ appends is

@@

Finally, appending the writes made for every item in [append], I finally find an explicit formula for $\nwritesfn(\varnitems)$.

@@

I wish to find the space allocated while appending $\varnitems$ items to a dynamic array. I denote this quantity $\spacefn(\varnitems)$, and define it as the total length of all arrays allocated during $\varnitems$ [append] calls. I now derive a formula for $\spacefn(\varnitems)$.

>>>>

\begin{description}
	\item[initial capacity] The capacity of a dynamic a array when one item is appended to it. I denote this $\initcapacity$.\\
	(The capacity of any dynamic array with size zero should be zero.)
	\item[growth factor] The factor by which the current capacity is multiplied to get the new capacity when $\listname$ grows. I denote this with $\growthfactor$.
\end{description}

\begin{algorithm}
	\begin{algorithmic}[1]
		\Procedure{$\FuncAppend$}{$L, item$}
			\If{$\listname.\FieldFull$}
				\State $\listname.\FuncGrow()$
			\EndIf

			\State $\listname.\FieldHead[\listname.\FieldHeadSize] \gets item$
			\State $\listname.\FieldSize \gets \listname.\FieldSize + 1$
		\EndProcedure
		\Statex
		\Procedure{$\FuncGrow$}{$L$}
			\If{$\listname.\FieldEmpty$}
				\State $\listname.\FieldBuffer \gets \FuncNewArray(\initcapacity)$
			\Else
				\State $new\ buf \gets \FuncNewArray(\listname.\FieldSize \cdot \growthfactor)$
				\State $\FuncArrayCopy(source: \listname.\FieldBuffer, dest: new\ buf, len: \listname.\FieldSize)$
				\State $\listname.\FieldBuffer \gets new\ buf$
			\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

Before I analyze the time complexity of $\FuncAppend$, I consider a different method for measuring its cost. Suppose I start with an empty list and $\varnitems$ elements are appended. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $\varnitems$ appends, and denote it $\nwritesfn(n)$.

In the above code, one write is performed after the conditional, so it is apparent that $\nwritesfn(\varnitems) \geq \varnitems$ after $n$ appends. However, $\FuncGrow$ also does some writing, so in order to find a precise formula for $\nwritesfn(\varnitems)$, I need to analyze when $\FuncGrow$ is called. Consider the following lemma:

@@

[proof]
The capacity starts out at zero and becomes $\initcapacity$ once one element is appended. From then on, it can only grow by a factor of $\growthfactor$ via $\FuncGrow$, so if $\growthfactor^i\initcapacity$ is the current capacity then $\growthfactor^{i + 1}\initcapacity$ must be the next capacity. By induction, the rest of the sequence is $$\left\{ \growthfactor^i\initcapacity \right\}_{i = 1}^k$$ for some $k$. Since the capacity is only as big as is needs to be, namely greater than or equal to $\varnitems$, $k$ is the smallest integer for which $\growthfactor^k\initcapacity \geq \varnitems$. Using this property, I now derive $k$.

@@

[make lemma]
Now, consider that to grow by $\growthfactor$, $\FuncGrow$ must be called. So $\growthfactor^k\initcapacity$ being present in $S_C$ means $\FuncGrow$ was called at size $\growthfactor^{k - 1}\initcapacity$ for $k \geq 1$. Then $\FuncGrow$ was called at the sizes $$S_R = 0,\ \initcapacity,\ \growthfactor\initcapacity,\ \growthfactor^2\initcapacity,\ \ldots\ \growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil - 1}\initcapacity$$

Each time $\FuncGrow$ is called, $i$ items are copied where $i$ is the current size of the dynamic array. Then the total number of items copied for $\varnitems$ appends is

@@

Finally, appending the writes made for every item in $\FuncAppend$, I finally find an explicit formula for $\nwritesfn(\varnitems)$.

@@

I wish to find the space allocated while appending $\varnitems$ items to a dynamic array. I denote this quantity $\spacefn(\varnitems)$, and define it as the total length of all arrays allocated during $\varnitems$ $\FuncAppend$ calls. I now derive a formula for $\spacefn(\varnitems)$.

### 2017-10-18 Draft2/Appending.Growth.tex
Rename some things and remove some brackets

[That lemma] still holds, since all of the assumptions made there are also true for growth arrays. In particular, although growth arrays resize slightly differently than dynamic arrays, the following claim is still true:

@@

For the base case, consider the second time [resize] is called: the current capacity is $\initcapacitynew$, and the next (total) capacity should be $\growthfactornew\initcapacitynew$. Then a buffer of size $(\growthfactornew - 1)\initcapacitynew$ should be appended, and that is what the algorithm does.

@@

Now I take a look at [resize]. Unlike dynamic arrays, the number of writes performed does not equal $i$ if the current size is $i$. No items are being copied; the only source of writes is appending a buffer to the tail. The goal is to find the number of writes that causes.

Let $\nwritesresizefn(n)$ denote the total number of writes made by [resize]. From [make lemma], I know that [resize] is called at the sizes

@@

Let $k$ be the number of times the tail is appended to. Since the tail is a dynamic array, the number of writes it makes is $\nwritesfn(k)$. Since [resize] is called $|S_R|$ times and it appends to the tail each time except for the first, $k = |S_R| - 1 = \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil$. Then the formula for $\nwritesresizefn(n)$ is

$$
\nwritesresizefn(n) = k + \left( \frac{\growthfactor^{\lceil \log_{\growthfactor} k - \log_{\growthfactor} \initcapacity \rceil} - 1}{\growthfactor - 1} \right) \initcapacity
$$

@@

% Should this be C(n) instead of C?
The space function for $\varnitems$ [append] calls on a growth array is denoted $\spacenewfn(\varnitems)$. Since growth arrays never throw away item buffers, if the current capacity is $C$ then the total length of item buffers is also $C$. $\spacenewfn(\varnitems)$ is usually greater than $C$, however; this is because growth arrays not only allocate item buffers, but stores references to them in the tail. Thus the space the tail allocates must also be determined.

>>>>

[That lemma] still holds, since all of the assumptions made there are also true for growth arrays. In particular, although growth arrays grow differently than dynamic arrays, the following claim is still true:

@@

For the base case, consider the second time $\FuncGrow$ is called: the current capacity is $\initcapacitynew$, and the next (total) capacity should be $\growthfactornew\initcapacitynew$. Then a buffer of size $(\growthfactornew - 1)\initcapacitynew$ should be appended, and that is what the algorithm does.

@@

Now I take a look at $\FuncGrow$. Unlike dynamic arrays, the number of writes performed does not equal $i$ if the current size is $i$. No items are being copied; the only source of writes is appending a buffer to the tail. The goal is to find the number of writes that causes.

Let $\nwritesgrowfn(n)$ denote the total number of writes made by $\FuncGrow$. From [make lemma], I know that $\FuncGrow$ is called at the sizes

@@

Let $k$ be the number of times the tail is appended to. Since the tail is a dynamic array, the number of writes it makes is $\nwritesfn(k)$. Since $\FuncGrow$ is called $|S_R|$ times and it appends to the tail each time except for the first, $k = |S_R| - 1 = \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil$. Then the formula for $\nwritesgrowfn(n)$ is

$$
\nwritesgrowfn(n) = k + \left( \frac{\growthfactor^{\lceil \log_{\growthfactor} k - \log_{\growthfactor} \initcapacity \rceil} - 1}{\growthfactor - 1} \right) \initcapacity
$$

@@

% Should this be C(n) instead of C?
The space function for $\varnitems$ $\FuncAppend$ calls on a growth array is denoted $\spacenewfn(\varnitems)$. Since growth arrays never throw away item buffers, if the current capacity is $C$ then the total length of item buffers is also $C$. $\spacenewfn(\varnitems)$ is usually greater than $C$, however; this is because growth arrays not only allocate item buffers, but stores references to them in the tail. Thus the space the tail allocates must also be determined.
